Structure Learning for Causal Decision Making

### 2019-03-03 UAI_2019/UAI_Draft.tex
cambiando intro y related work

\section{Introduction}
Decision making under uncertain conditions is a fundamental task for intelligent reasoning and a any day task for human beings and organizations (\cite{savage1954the}, \cite{danks2014unifying}, \cite{lake2017building}). It is a well known fact that if an agent has \textit{rational preferences} then he must choose according to the Maximum Expected Utility criteria, in which the agent chooses an action by considering both his preferences, which are represented by an utility function, and the probability of occurrence of outcomes: with respect to this point, the agent either knows the probabilities in the environment, or, if he does not, by considering a \textit{subjective} probability measure (\cite{von1944theory}, \cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}). We consider the case where the uncertainty component in the decision problem is controlled by a \textit{causal mechanism}, represented as a Causal Graphical Model which is always unknown by the agent and assumed to remain fixed and invariant to interventions (\cite{woodward2005making}). It has been proposed a mechanism for a decision maker to find good actions by holding beliefs about the causal relations in the environment, using them to make an action and then updating those beliefs in terms of what is caused by a chosen action. Such mechanism, which is described in \cite{gonzalez2018playing} is limited since it is assumed that the decision maker knows the \textit{structure} of the Causal Graphical Model. Here we attempt to relax such assumption and allowint the decision maker to know only the variables in the model, with the help of an \textit{ordering} in the variables in such a way that it is know which variables can not cause each other.

\section{On Causality and Learning}
Several definitions of Causality have been attempted by philosophers and scientists, such as those described in \cite{holland1986statistics}. We use here the formal definition of Causality given by \cite{spirtes2000causation}, which belongs to the family of \textit{manipulationist} theories as described by \cite{woodward2005making}. We consider Causal Graphical Models (\cite{koller2009probabilistic}) as representations of the causal relations that hold in an environment and the resulting interventional distribution.

@@

\section{Related Work}
In \cite{lattimoreNIPS2016}, the problem of finding a best intervention over a known causal model for a fixed number of learning trials is considered, their assumption of knowing the causal moder is later relaxed by \cite{sen2017identifying} who considers interventions over a partially unknown causal model. By following \cite{joyce1999foundations} intuitions, in \cite{gonzalez2018playing} it is proposed that a rational decision maker can find an optimal action by holding \textit{beliefs} about the existence and strength of causal relations and by using those beliefs \textit{as if} they were the true relations in order to make a choice and then updating those beliefs after observing what the chosen action \textit{caused} good actions can be found. Such procedure implicitly considers a distribution over Causal Graphical Models, where each structure is used locally (in each step) to make a choice. In \cite{2019arXiv190202279G}, such authors propose a formal criteria which has the form:
\[  a^\ast = \textrm{ argmax }_{a \in \mathcal{A}}\sum_g P(Y=1 | do(a), g)P(g), \]
where $\mathcal{A}$ is the set of available actions to the decision maker and $1$ is the most desired value of the \textit{objective variable} $Y$, for the decision maker. We notice that since the action chosen is the action with the highest probability of causally producing the most desired action, then $a^\ast$ necesarilly is the action that maximizes the expected utility for the agent. The function above described is indeed a counterfactual distribution, since it allows to compare the probability of causing a desired actions by alternative actions that were not chosen.

While \cite{pearl2009causality} proposes that a rational agent in the presence of a causal environment should optimize the function $U(x)$ defined as 
\[ U(x)=\sum_y P(y | do(x)) u(y), \]
we notice that this function does not consider the agent's beliefs, or current knowledge, about the causal structure that holds in his environment. We note that the criteria proposed by \cite{2019arXiv190202279G} makes use of an implicit probability distribution over causal structures, since the decision maker holds beliefs about the causal nature of his environment and within each realization of those beliefs the maximization for the optimal action is performed, while Pearl's function assumes a single causal structure.

Using a version of the criteria described above, in \cite{gonzalez2018playing} a decision-making procedure is proposed for the case when the agent knows the \textit{structure} of the Causal Model which controls his environment. It remains an open question to tackle the problem of a fully unknown structure. It is known that from observational data alone, a Causal Graphical Model is identifiable up to Markov equivalence without any the assumption of a specific functional model with given error distributions (\cite{hauser2012characterization}, \cite{hauser2012two}, \cite{peters2011identifiability}). We pretend to extend here the work of \cite{gonzalez2018playing} to the case of a rational decision maker who knows the variables that control his environment and when he has extra knowledge of variables that can not be cause of each other.

>>>>

\section{Introduction}
Decision making under uncertain conditions is a fundamental task for intelligent reasoning and an any-day task for human beings and organizations (\cite{savage1954the}, \cite{danks2014unifying}, \cite{lake2017building}). It is a well known fact that if an agent has \textit{rational preferences} then he must choose according to the Maximum Expected Utility criteria, in which the agent chooses an action by considering both his preferences, which are represented by an utility function, and the probability of occurrence of outcomes: with respect to this point, the agent either knows the probabilities in the environment, or, if he does not, by considering a \textit{subjective} probability measure (\cite{von1944theory}, \cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}). The case where the uncertainty component in the decision problem is controlled by a \textit{causal mechanism}, has been previously studied in \cite{joyce1999foundations}, \cite{pearl2009causality}, \cite{lattimoreNIPS2016}, \cite{sen2017identifying}, \cite{gonzalez2018playing}, \cite{2019arXiv190202279G}. In particular, \cite{joyce1999foundations} provided the intuition that a decision maker should consider what is \textit{caused} by his actions, and \cite{pearl2009causality} proposed decision-making by maximizing a function involving the well known \textit{do} operator.

In this work we consider that the causal mechanism controlling the environment is represented as a Causal Graphical Model which is unknown by the agent and assumed to remain fixed and invariant to interventions (\cite{woodward2005making}). In recent papers, such as those by \cite{lattimoreNIPS2016}, \cite{sen2017identifying} and \cite{gonzalez2018playing} it is assumed that the decision maker knows either all of, or a certain aspect such as the structure, of the causal model. Here we attempt to relax such assumption and allowint the decision maker to know only the variables in the model, with the help of an \textit{ordering} in the variables in such a way that it is know which variables can not cause each other.

\section{On Causality and Learning}
Several definitions of Causality have been attempted by philosophers and scientists, such as those described in \cite{holland1986statistics}. We use here the formal definition of Causality given by \cite{spirtes2000causation}, which belongs to the family of \textit{manipulationist} theories as described by \cite{woodward2005making}. We consider Causal Graphical Models (\cite{koller2009probabilistic}, \cite{sucar2015probabilistic}) as representations of the causal relations that hold in an environment and the resulting interventional distribution.

@@

\section{Related Work}
In \cite{lattimoreNIPS2016}, the problem of finding a best intervention over a known causal model for a fixed number of learning trials is considered and it was found that by considering causal information a \textit{faster} optimal-action learning can be achieved. Their assumption of knowing the causal model was relaxed by \cite{sen2017identifying} who considers a partially known causal model and allows interventions over a the unknown part of the model. In \cite{gonzalez2018playing} it was proposed that a rational decision maker can find an optimal action by holding \textit{beliefs} about the existence and strength of causal relations and by using those beliefs \textit{as if} they were the true relations in order to make a choice and then updating those beliefs after observing what the chosen action \textit{caused} good actions can be found. Such procedure implicitly considers a distribution over Causal Graphical Models, where each structure is used within each step in otder to make a choice by considering his causal knowledge as it if were true, as proposed by \cite{joyce1999foundations}. Their work assumes that the decision maker knows the structure, but not the associated parameters, of the model. Later, \cite{2019arXiv190202279G}, such authors formalize that procedure by stating a formal criteria which has the form:
\[  a^\ast = \textrm{ argmax }_{a \in \mathcal{A}}\sum_g P(Y=1 | do(a), g)P(g), \]
where $\mathcal{A}$ is the set of available actions to the decision maker and $1$ is the most desired value of the \textit{objective variable} $Y$, for the decision maker. We notice that since the action chosen is the action with the highest probability of causally producing the most desired action, then $a^\ast$ necesarilly is the action that maximizes the expected utility for the agent. The function above described is indeed a counterfactual distribution, since it allows to compare the probability of causing a desired actions by alternative actions that were not chosen.

While \cite{pearl2009causality} proposes that a rational agent in the presence of a causal environment should optimize the function $U(x)$ defined as 
\[ U(x)=\sum_y P(y | do(x)) u(y), \]
we notice that this function does not consider the agent's beliefs, or current knowledge, about the causal structure that holds in his environment. Pearl's function assumes a single causal structure. as opposed by \cite{2019arXiv190202279G} who make use of an implicit probability distribution over causal structures.

It remains an open question to tackle the problem of a fully unknown structure. It is known that from observational data alone, a Causal Graphical Model is identifiable up to Markov equivalence without any the assumption of a specific functional model with given error distributions (\cite{hauser2012characterization}, \cite{hauser2012two}, \cite{peters2011identifiability}). We pretend to relax the mentioned limitations to the case of a rational decision maker who knows the variables that control his environment and he has extra knowledge of variables that, in principle can not be cause of each other.
